\begin{center}
{\large Ed Summers}\\
ehs@pobox.com / 240-478-7086\\
304 Brewster Court, Silver Spring, MD 20901\\
\end{center}

I am a software developer and research designer with two decades of
experience bridging the worlds of data management and the Web. I have
worked in academia, startups, corporations and government. I thrive
while working in agile, highly collaborative teams that focus on
knowledge sharing and a commitment to sustainability.

\hypertarget{skills}{%
\subsection{Skills}\label{skills}}

\begin{itemize}
\tightlist
\item
  web technology: APIs, REST, linked data, web archiving
\item
  languages: Python, JavaScript, Ruby, PHP, Perl, Go, Java, C, Scheme
\item
  data analysis: Pandas, SciPy, Spark, Presto
\item
  infrastructure: Linux, AWS, Docker, Ansible, Terraform
\item
  databases: PostgreSQL, MariaDB, Redis, ElasticSearch, Solr, MongoDB
\item
  visualization: D3, Plotly, vis.js, Gephi
\item
  practice: agile, automated testing, continuous integration, standards,
  documentation
\end{itemize}

\hypertarget{work-history}{%
\subsection{Work History}\label{work-history}}

Senior Software Developer\\
Digital Library Systems and Services\\
Stanford University\\
01/2022 -

\begin{itemize}
\tightlist
\item
  Full stack software development (Ruby, Python, JavaScript)
\item
  Web archive creation, preservation and access
\item
  Digital repositories, data preservation
\item
  Web services (REST) / Service design
\end{itemize}

Design Researcher\\
Webrecorder\\
10/2021 - 02/2023

\begin{itemize}
\tightlist
\item
  Requirements and use case gathering
\item
  Technical writing for data
  \href{https://specs.webrecorder.net}{specifications}
\item
  Software development (Python, JavaScript)
\end{itemize}

Research Faculty\\
Maryland Institute for Technology in the Humanities\\
University of Maryland\\
08/2014 - 12/2021

\begin{itemize}
\tightlist
\item
  Full stack software development (JavaScript, React, Python,
  PostgreSQL, ElasticSearch)
\item
  Social media data analysis and visualization (Python, JavaScript, D3,
  VisJS)
\item
  Linked Data publishing (Python, NodeJS)
\item
  Sustainable data preservation workflows (AWS)
\item
  Automated testing and continuous integration (GitHub)
\item
  Configuration management (Git, Docker, Ansible)
\item
  Agile requirements gathering \& project management
\item
  Instructor \& Mentor: software development (Python, JavaScript), data
  science (Python, Pandas, Spark)
\end{itemize}

Consultant\\
The Brookings Institution\\
Washington, DC\\
June 2021 - September 2021

\begin{itemize}
\tightlist
\item
  Social media data collection and ETL (Twitter, Reddit, Python)
\item
  Network analysis and visualization (Python, React, vis.js)
\item
  Report writing and research data publishing
\end{itemize}

Visiting Researcher\\
National Institute of Standards and Technology\\
Gaithersburg, Maryland\\
August 2018 - August 2021

\begin{itemize}
\tightlist
\item
  Conducted a multi-year field study to analyze data acquisition methods
  used in building the world's largest public archive of computer
  software (NSRL).
\item
  Software development and research design for data data preservation
  packaging formats (Python, Git, Jupyter, PostgreSQL, BagIt, DFXML,
  OCFL)
\end{itemize}

Information Technology Specialist\\
Library of Congress\\
Washington, DC\\
03/2006 - 08/2014

\begin{itemize}
\tightlist
\item
  Digital preservation and repository software design and development
  (Python, Java, MySQL)
\item
  Lead software developer on the
  \href{http://chroniclingamerica.loc.gov/}{National Digital Newspaper
  Project} (Django, MySQL)
\item
  Created a web based quality assurance tool for handling high volume
  image digitization projects
\item
  Social media archiving (Twitter)
\item
  Designed and implemented the \href{http://id.loc.gov/}{Library of
  Congress Linked Data Service}
\item
  Specification writing and testing
  \href{https://www.w3.org/2006/07/SWD/}{W3C}
\end{itemize}

Consultant\\
George Washington University\\
Washington, DC\\
10/2013 - 12/2014

\begin{itemize}
\tightlist
\item
  Django Web Development
\item
  Voyager ILS integration (Oracle)
\item
  Serial Solutions Summon API integration
\end{itemize}

Consultant\\
Nelson Institute for Environmental Studies\\
Madison, WI\\
09/2013 - 06/2014

\begin{itemize}
\tightlist
\item
  Web data harvesting (Python)
\item
  Data packaging and publishing (Git, AWS)
\end{itemize}

Consultant\\
Hypothes.is\\
San Francisco, CA\\
01/2013 - 10/2013

\begin{itemize}
\tightlist
\item
  JavaScript and Python development
\item
  Development of the open source
  \href{http://github.com/okfn/anotator}{Annotator} and
  \href{http://github.com/okfn/anotator-store}{Annotator Store} projects
\item
  Design and implementation of {[}cross format annotation{]} annotation
  support
\item
  Functional/unit testing (Selenium, Pytest, Travis-CI, SauceLabs)
\end{itemize}

Consultant\\
GlueJar\\
Montclair, NJ 11/2011 - 04/2012

\begin{itemize}
\tightlist
\item
  Django software development
\item
  AWS (EC2, RDS, Route 53)
\item
  Google Books API harvesting
\item
  Google, Facebook, Twitter social media integration
\item
  Automated testing and continuous integration (Jenkins)
\end{itemize}

Consultant\\
Yale School of Medicine\\
New Haven, CT\\
07/2005 - 12/2005

\begin{itemize}
\tightlist
\item
  Python object oriented design
\item
  Automated testing
\item
  Implemented a reporting framework
\item
  Quixote, Pylucene, MySQL, Red Hat administration
\end{itemize}

Senior Software Engineer\\
Follett Corporation\\
McHenry IL\\
07/2002 - 03/2006

\begin{itemize}
\tightlist
\item
  Web service design and implementation
\item
  Java, PHP and Perl software development
\item
  Automated testing and performance tuning
\item
  Database administration (Oracle, SQL Server)
\item
  XML metadata harvesting and ingest
\item
  Requirements gathering and project planning
\end{itemize}

Consultant\\
Performix Chicago IL\\
02/2004 - 04/2005

\begin{itemize}
\tightlist
\item
  Object oriented design
\item
  Refactored large Perl codebase
\item
  Perl, Apache, mod\_perl, DB2
\end{itemize}

Consultant\\
Queens Borough Public Library\\
New York, NY\\
06/2004 - 05/2005

\begin{itemize}
\tightlist
\item
  Character encoding conversion
\item
  Controlled vocabulary mining
\item
  Perl data processing training
\end{itemize}

Consultant\\
Notre Dame University\\
South Bend IN\\
06/2004 - 12/2004

\begin{itemize}
\tightlist
\item
  Training in object oriented programming and unit testing
\item
  High volume metadata harvesting
\item
  Guidance and implementation search technologies
\item
  Ported Java Common Query Language parser to Perl
\end{itemize}

Senior Software Engineer\\
CheetahMail\\
New York NY\\
02/2000 - 07/2002

\begin{itemize}
\tightlist
\item
  Team lead in a group of 8 programmers
\item
  Developed data processing infrastructure for Fortune 500 clients
\item
  Designed/developed real time reporting system
\item
  Creation and maintenance of web service API
\item
  Expertise in SMTP and HTTP protocols
\item
  Web services development
\item
  Technical representative for partners and sales
\item
  Networking and systems administration as required
\end{itemize}

Metadata Specialist\\
Old Dominion University\\
Norfolk, VA\\
04/1998 - 02/2000

\begin{itemize}
\tightlist
\item
  SGML and XML data processing for the VIVA Consortium
\item
  Created current awareness system to alert students/faculty to new
  items in their field of interest.
\item
  ILS and OCLC metadata workflow management
\end{itemize}

Visiting Assistant Professor University of Illinois\\
Champaign-Urbana, IL\\
08/1996 - 03/1998

\begin{itemize}
\tightlist
\item
  Web development (ASP, SQL Server)
\item
  Collection development and cataloging in Sociology and Political
  Science
\item
  Preservation of 19th century criminology pamphlet collection
\item
  Managed reference services in the Social Science Library
\end{itemize}

\hypertarget{selected-projects}{%
\subsection{Selected Projects}\label{selected-projects}}

These are some projects and experiments where I've been directly
involved in the technical development.

\begin{itemize}
\tightlist
\item
  \href{https://library.stanford.edu/projects/web-archiving}{Stanford
  Web Archive Program} - On site crawling, preservation and replay of
  web archives using \href{https://webrecorder.net}{Webrecorder} tools.
\item
  \href{https://unlockingtheairwaves.org}{Unlocking the Airwaves} - NEH
  funded virtual reunification and minimal computing project
\item
  \href{https://www.docnow.io}{Documenting the Now} - Mellon funded
  social media archiving practices and tools
\item
  \href{https://chroniclingamerica.loc.gov}{Chronicling America} -
  Library of Congress access to historical newspapers.
\item
  \href{https://id.loc.gov}{Linked Data Service} - Publishing controlled
  vocabularies at the Library of Congress using Linked Data
\item
  \href{https://web.hypothes.is/}{Hypothesis} - An annotation tool for
  the web
\item
  \href{http://wikistream.wmflabs.org}{Wikistream} - visualize realtime
  edits to all major language Wikipedias
\item
  \href{http://github.com/edsu/anon}{Anon} - track anonymous Wikipedia
  edits using Twitter
\item
  \href{http://github.com/edsu/pymarc}{Pymarc} - a python library for
  MARC processing
\item
  \href{https://jobs.code4lib.org}{jobs.code4lib.org} - a jobs board for
  the code4lib community
\item
  \href{http://github.com/edsu/linkypedia}{Linkypedia} - view how
  websites are used on Wikipedia
\item
  \href{http://github.com/fondz}{Fondz} - generate an archival
  description for describing born digital content
\item
  \href{https://github.com/edsu/microdata}{Microdata} - a python library
  for parsing HTML 5 microdata
\item
  \href{https://github.com/RDFLib}{rdflib} - a Python library for
  processing RDF
\end{itemize}

\hypertarget{selected-publications}{%
\subsection{Selected Publications}\label{selected-publications}}

\begin{itemize}
\tightlist
\item
  \href{https://manifold.as.uky.edu/read/teaching-demagoguery-and-democracy-rhetorical-pedagogy-in-polarized-times/section/a2669e5b-6f16-4422-b384-2f859bd88605}{Demagoguery,
  Democracy, and Digitality}. In Teaching Demagoguery and Democracy:
  Rhetorical Pedagogy in Polarized Times. November, 2021.
\item
  \href{https://www.brookings.edu/research/bystander-intervention-on-social-media-examining-cyberbullying-and-reactions-to-systemic-racism/}{Bystander
  intervention on social media: Examining cyberbullying and reactions to
  systemic racism}. The Brookings Institution. October, 2021.
\item
  \href{https://drum.lib.umd.edu/handle/1903/26731}{Legibility Machines:
  Archival Appraisal and the Genealogies of Use}. PhD Dissertation.
  University of Maryland. December, 2020.
\item
  \href{https://inkdroid.org/papers/appraisal-talk.pdf}{Appraisal Talk
  in Web Archives}. Archivaria. May, 2020.
\item
  \href{https://doi.org/10.17723/0360-9081-83.1.167}{What We Talk about
  When We Talk about Archiving the Web}. American Archivist. March,
  2020.
\item
  \href{https://culturalanalytics.org/article/11048}{Archival
  Circulation on the Web: The Vine-Tweets Dataset}. Cultural Analytics.
  June, 2019.
\item
  \href{https://www.tandfonline.com/doi/abs/10.1080/1369118X.2019.1626469}{Algorithms
  and Agenda-Setting in Wikileaks' \#PodestaEmails Release}.
  Information, Communication \& Society. 2019
\item
  \href{https://osf.io/preprints/socarxiv/75mjp/}{Appraisal Practices in
  Web Archives}. SocArXiv Papers, 2018.
\item
  \href{https://www.tandfonline.com/doi/abs/10.1080/24701475.2017.1383639}{Twitter
  and Tear Gas: Book Review}. Internet Histories, 2017.
\item
  \href{http://dx.doi.org/10.1080/01419870.2017.1335422}{Ferguson and
  the death of Michael Brown on Twitter: \#BlackLivesMatter, \#TCOT, and
  the evolution of collective identities}. Ethnic and Racial Studies,
  40(10).
\item
  \href{http://dx.doi.org/10.1080/01419870.2017.1334934}{\#sayhername: a
  case study of intersectional social media activism}. Ethnic and Racial
  Studies, 40(10).
\item
  \href{/papers/bots-seeds-people.pdf}{Bots, Seeds and People: Web
  Archives as Infrastructure}. Proceedings of the 2017 ACM Conference on
  Computer Supported Cooperative Work and Social Computing, CSCW 17.
\item
  \href{https://source.opennews.org/articles/automating-transparency/}{Automating
  Transparency}. The Source. April, 2015.
\item
  \href{http://arxiv.org/abs/1303.4702}{MJ no more: Using Concurrent
  Wikipedia Edit Spikes with Social Network Plausibility Checks for
  Breaking News Detection}. 2013.
\item
  \href{http://arxiv.org/abs/1302.4591}{Linking Things on the Web: A
  Pragmatic Examination of Linked Data}, 2013.
\item
  \href{http://arxiv.org/abs/1302.1224}{Key Choices in the Design of
  Simple Knowledge Organization System (SKOS)}, 2013.
\item
  \href{http://www.w3.org/2005/Incubator/lld/XGR-lld-20111025/}{Library
  Linked Data Final Report}, W3C Incubator Group Report, October 2011.
\item
  \href{http://www.w3.org/TR/skos-primer}{SKOS Simple Knowledge
  Organization System Primer}, W3C Working Group Note, August 18th,
  2008. http://www.w3.org/TR/skos-primer/
\item
  \href{http://dcpapers.dublincore.org/pubs/article/view/916}{LCSH, SKOS
  and Linked Data}, Proceedings of the International Conference on
  DublinCore and Metadata Applications, 2008.
\item
  \href{http://inkdroid.org/journal/2008/01/04/following-your-nose-to-the-web-of-data/}{Following
  Your Nose to the Web of Data}, Information Standards Quarterly, 2008
\item
  \href{http://www.ariadne.ac.uk/issue38/summers}{Building OAI-PMH
  Harvesters With Net::OAI::Harvester} Ariadne, no. 38, Jan 2004.
\item
  \href{https://dx.doi.org/10.1108/07378830410524459}{Cyril: Expanding
  the Horizons of MARC21}, Library Hi Tech, v. 22, no. 1, 2004.
\item
  \href{https://dx.doi.org/10.1300/J121v16n01_04}{A Consortial Approach
  to Cooperative Cataloging and Authority Control: The Virtual Library
  of Virginia Experience}, Resource Sharing and Information Networks, v.
  16, no. 1, 2003, p.~33-53.
\item
  \href{https://web.archive.org/web/20130625070202/https://www.theperlreview.com/articles/php.html}{Paying
  Homage to Perl (PHP)}, The Perl Review, January, 2003.
\item
  MARC it your way: MARC.pm, Information Technology and Libraries,
  March, 2002.
\item
  \href{https://doi.org/10.1016/S1464-9055(99)00096-2}{Comprehensive
  Access to Periodicals: A Database Solution}, Library Collections
  Acquisitions and Technical Services, v. 24, no. 2, 2000, p.~33-44.
\item
  \href{https://doi.org/10.7771/2380-176X.3746}{Chaos--Show Me the
  Metadata}, Against the Grain, v.11, no.1, 1998.
\item
  \href{https://crln.acrl.org/index.php/crlnews/article/view/18375/20725}{Gateways
  to Social Work/Welfare on the Net}, C\&RL News, v.59, no. 3, 1998,
  p.163-167.
\end{itemize}

\hypertarget{education}{%
\subsection{Education}\label{education}}

\begin{itemize}
\tightlist
\item
  PhD, College of Information Studies, University of Maryland,
  2015-2020.
\item
  MLS, School of Communications, Information and Library Science,
  Rutgers University, 1995-1996.
\item
  BA English and American Literature, Rutgers University, 1987-1991.
\end{itemize}

\hypertarget{awards}{%
\subsection{Awards}\label{awards}}

\begin{itemize}
\tightlist
\item
  2018
  \href{https://www.diglib.org/announcing-the-2018-dlf-community-capacity-award-winner/}{Community
  Capacity Award}. Digital Library Federation.
\item
  2018 \href{https://www2.archivists.org/node/23172}{Exemplary Service
  Award}. Society of American Archivists.
\item
  2015 University of Maryland Scholarship and Research Award
\item
  2015
  \href{http://www.ala.org/news/press-releases/2015/03/ed-summers-recipient-2015-kilgour-award}{Frederick
  G. Kilgour Award for Research in Library and Information Technology}.
  American Library Association.
\end{itemize}
