\section{Ed Summers}

Silver Spring, Maryland, USA

\href{mailto:ehs@pobox.com}{\texttt{ehs@pobox.com}}\\\url{http://inkdroid.org}\\\url{http://twitter.com/edsu}\\\url{http://github.com/edsu}

\subsection{Expertise}

\begin{itemize}
\item
  web technology, REST, linked data\\
\item
  digital preservation, metadata, search\\
\item
  opensource, standards development, community building\\
\item
  automated testing, agile software development, prototyping\\
\item
  python, javascript, ruby, perl, php, java, c, scheme
\end{itemize}
\subsection{Work History}

Information Technology Specialist\\Repository Development
Center\\Library of Congress, Washington, DC\\03/2006 - present

\begin{itemize}
\item
  digital preservation and repository software design and development
\item
  lead software developer on the
  \href{http://chroniclingamerica.loc.gov/}{National Digital Newspaper Project}
\item
  created a web based quality assurance tool for handling high volume
  image digitization projects
\item
  digital archiving of social media (Twitter)
\item
  designed and implemented the
  \href{http://id.loc.gov/}{Library of Congress Linked Data Service}
\item
  contributing member to the Semantic Web Deployment Group, and
  Library Linked Data Incubator Group at the W3C
\end{itemize}
Consultant\\Canary Project\\Yale School of Medicine, New Haven,
CT\\07/2005 - 12/2005

\begin{itemize}
\item
  Python object oriented design
\item
  automated testing
\item
  implemented a reporting framework
\item
  quixote, pylucene, mysql, redhat administration
\end{itemize}
Senior Software Engineer\\Follett Corporation, McHenry IL\\05/2005
- 03/2006

\begin{itemize}
\item
  building and supporting a distributed knowledge management
  application
\item
  requirements gathering and project planning
\item
  web service design and implementation
\item
  ONIX metadata harvesting and ingest
\item
  Java performance tuning
\item
  automated testing
\item
  database administration
\item
  Linux/Windows systems administration
\end{itemize}
Consultant\\Performix, Chicago IL\\02/2004 - 04/2005

\begin{itemize}
\item
  object oriented design using CRC cards
\item
  refactored large Perl codebase
\item
  Perl, Apache, mod\_perl, DB2
\end{itemize}
Consultant\\Queens Borough Public Library, New York NY\\06/2004 -
05/2005

\begin{itemize}
\item
  character encoding conversion
\item
  controlled vocabulary mining
\item
  Perl data processing training
\end{itemize}
Consultant\\Notre Dame University, South Bend IN 06/2004 - 12/2004

\begin{itemize}
\item
  training in object oriented programming and unit testing
\item
  high volume metadata harvesting
\item
  guidance and implementation search technologies
\item
  ported Java Commonn Query Language parser to Perl
\end{itemize}
Applications Developer\\Follett Library Resources, McHenry
IL\\07/2002 - 05/2004

\begin{itemize}
\item
  PHP/Perl development in a Unix/Oracle environment
\item
  determined project requirements, created task lists and schedules
\item
  designed and developed library collection analysis application
\item
  unit and application level automated testing
\item
  Oracle/Apache stress testing and performance tuning
\item
  XML processing
\item
  distributed application design using SOAP/XMLRPC
\end{itemize}
Senior Software Engineer CheetahMail, New York NY 02/2000 -
07/2002

\begin{itemize}
\item
  team lead in a group of 8 programmers
\item
  developed data processing infrastructure for Fortune 500 clients
\item
  designed/developed real time reporting system
\item
  creation and maintainance of web service API
\item
  in depth knowledge of SMTP and HTTP protocols
\item
  web services development
\item
  technical representative for partners and sales
\item
  networking and systems administration as required
\end{itemize}
Electronic Resources Librarian\\Old Dominion University, Norfolk,
VA\\04/1998 - 02/2000

\begin{itemize}
\item
  SGML data processing for electronic texts for 39 universities in
  Virginia.
\item
  created current awareness system to alert students/faculty to new
  items in their field of interest.
\item
  created XML publishing system for archival finding aids.
\item
  developed serials database for locating licensed fulltext.
\item
  III/OCLC metadata management integration
\end{itemize}
Visiting Assistant Professor\\University of Illinois,
Champaign-Urbana, IL\\08/1996 - 03/1998

\begin{itemize}
\item
  web/database development (ASP/SQLServer)
\item
  collection development and cataloging in Sociology and Political
  Science
\item
  preservation of 19th century criminology pamphlet collection
\item
  managed reference services in the Social Science Library
\end{itemize}
\subsection{Selected Publications}

\begin{itemize}
\item
  \href{http://www.w3.org/2005/Incubator/lld/XGR-lld-20111025/}{Library Linked Data Final Report},
  W3C Incubator Group Report, October 2011.
\item
  \href{http://www.w3.org/TR/skos-primer}{SKOS Simple Knowledge Organization System Primer},
  W3C Working Group Note, August 18th, 2008.
  http://www.w3.org/TR/skos-primer/
\item
  \href{http://dcpapers.dublincore.org/ojs/pubs/article/view/916}{LCSH, SKOS and Linked Data},
  Proceedings of the International Conference on DublinCore and
  Metadata Applications, 2008.
\item
  \href{http://inkdroid.org/journal/2008/01/04/following-your-nose-to-the-web-of-data/}{Following Your Nose to the Web of Data},
  Information Standards Quarterly, 2008
\item
  \href{http://www.ariadne.ac.uk/issue38/summers}{Building OAI-PMH Harvesters With Net::OAI::Harvester},
  Ariadne, no. 38, Jan 2004.
\item
  Cyril: Expanding the Horizons of MARC21, Library Hi Tech, v. 22,
  no. 1, 2004.
\item
  A Consortial Approach to Cooperative Cataloging and Authority
  Control: The Virtual Library of Virginia Experience, Resource
  Sharing and Information Networks, v. 16, no. 1, 2003, p.~33--53.
\item
  \href{http://www.theperlreview.com/articles/php.html}{Paying Homage to Perl (PHP)},
  The Perl Review, January, 2003.
\item
  MARC it your way: MARC.pm, Information Technology and Libraries,
  March, 2002.
\item
  Comprehensive Access to Periodicals: A Database Solution, Library
  Collections Acquisitions and Technical Services, v. 24, no. 2,
  2000, p.~33--44.
\item
  Show Me the Metadata, Against the Grain, v.11, no.1, 1998.
\item
  Gateways to Social Work/Welfare on the Net, C\&RL News, v.59, no.
  3, 1998, p.163--167.
\end{itemize}
\subsection{Education}

\begin{itemize}
\item
  MLS, School of Communications, Information and Library Science,
  Rutgers University, 1995--1996
\item
  BA English and American Literature, Rutgers University, 1987--1991.
\end{itemize}
